---
title: "Project #2"
author: "Rocco Matarazzo"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE) 

library(tidyverse)
library(rmarkdown)
library(ggplot2)
library(jsonlite)
library(httr)
library(ggcorrplot)

```

# Welcome to my API Vignette! 
Hello! This vignette is aimed at interacting with an API. Here we will interact with an MLB API. We will delve into multiple endpoints as well as some Exploratory Data Analysis (EDA). 

# Required Packages
The packages needed to create this markdown/GitHub post include tidyverse, rmarkdown, ggplot2, ggcorrplot, httr, and jsonlite.
The main packages that interact with the API itself are httr and jsonlite. The former four aid in the function creation for this document, the document itself, and the EDA.

# API Interaction Functions
Below are several different functions that pull from the MLB Data API. They work to acquire team and player information as well as player statistics.

`All Teams`  
This first interaction is rather simple, but useful! This will return all the teams for a given season.
```{r}

AllTeams <- function(year){
  
 ### This function is built to return each team for an MLB season ###
 ### Inputs:
 ### year -- a season of interest, i.e. 2023
 ### Example: AllTeams(2023) will return the 30 MLB teams who played in the 2023 season.
  
  
  # If a year is not given properly (i.e. not numeric) then 2023 will be the given value.
  if(is.numeric(year)){
    year <- year
  }else{
    year <- 2023
  }
  
  # First call to the API
  initialRead <- 
  GET(paste0("http://lookup-service-prod.mlb.com/json/named.team_all_season.bam?sport_code='mlb'&all_star_sw='N'&season=",year))

# We parse that data and get the data.frame object of interest
parsedJSON <- fromJSON(rawToChar(initialRead$content))
allTeams <- parsedJSON$team_all_season$queryResults$row

# Making the d.f. a tibble
Teams <- as_tibble(allTeams)
return(Teams)
}

```

`40-Man Roster for the 2023 Season`  
This second interaction will return a Major League Baseball's team's 40-man roster for the 2023 season. The team object should be entered in full format i.e. "Philadelphia Phillies" rather than "Phillies" or "Philadelphia." This is only available for the 2023 MLB season.

```{r}

FortyManRoster <- function(team){
 
 ### This function is built to return a team's 40-man roster for the 2023 MLB season ###
 ### Inputs:
 ### team -- a Major League Baseball team's full display name, i.e. "Philadelphia Phillies"
 ### Example: FortyManRoster("Philadelphia Phillies") returns the Phillies' 40-man roster for 2023

   # If a team is not given properly (i.e. not a characater) then the Phillies will be the given value.
  if(is.character(team)){
    team <- team
  }else{
    team <- "Philadelphia Phillies"
  }
  
# First we grab team-related information from the API for the given season
# using our very first function! 
# 40man roster is only available for 2023
allTeams <- AllTeams(2023)

# Here we select the individual teamId for the team of interest from the allTeams data.frame
teamId <- allTeams %>% filter(name_display_long == team) %>% select(team_id)

# We plug the teamId back into the API to return their 40-man roster
team <- GET(paste0("http://lookup-service-prod.mlb.com/json/named.roster_40.bam?team_id='",teamId,"'"))

# We parse the data, and ultimately return the 40-man roster
parsedTeam <- fromJSON(rawToChar(team$content)) 
FortyManFull <- parsedTeam$roster_40$queryResults$row

# Here we select columns of interest and rename them
# We also convert the d.f. into a tibble
as_tibble(FortyManFull %>% select(Team = team_name,
                        FullName = name_full,
                        FirstName = name_first,
                        LastName = name_last,
                        Position = position_txt,
                        Bats = bats,
                        Throws = throws,
                        Number = jersey_number,
                        PlayerID = player_id))
}

```

`All Players In a Season`  
This function will return every player who was on a MLB roster for a given season. I originally had this function wrapped within the Player Statistics' functions below. However, I figured it was a good idea to make it separate, and just use that within the Player Statistics' functions.
```{r}

LeagueRosters <- function(year){

 ### This function is built to return each MLB player in a given season ###
 ### Inputs:
 ### year -- a season of interest, i.e. 2023
 ### Example: LeagueRosters(2021) will return the rosters across the MLB in 2021

   # If a year is not given properly (i.e. not numeric) then 2023 will be the given value.
  if(is.numeric(year)){
    year <- year
  }else{
    year <- 2023
  }
  
# First we grab team-related information from the API for the given season
initialRead <- 
  GET(paste0("http://lookup-service-prod.mlb.com/json/named.team_all_season.bam?sport_code='mlb'&all_star_sw='N'&season=",year))

# We parse that data and get the data.frame object of interest
parsedJSON <- fromJSON(rawToChar(initialRead$content))
allTeams <- parsedJSON$team_all_season$queryResults$row
      
# Here we select each teamId and put it in a data frame
allTeamIds <- allTeams %>% select(team_id)
  
# We set a base object to be null
df <- NULL

# We create an inner-function called "rosterFunction" that we can eventually call for each team
    for(i in nrow(allTeamIds)){

     rosterFunction <- function(i){
       # We first get the JSON output for the given year for each team id
             team <- GET(paste0(
            "http://lookup-service-prod.mlb.com/json/named.roster_team_alltime.bam?start_season='"
            , year,"'&end_season='"
            , year,"'&team_id='"
            ,i,"'"))
            # We parse the data, and ultimately return the roster for that season
            parsedTeam <- fromJSON(rawToChar(team$content)) 
            RosterOuput <- parsedTeam$roster_team_alltime$queryResults$row
            # Here we select columns of interest and rename them
            roster <- RosterOuput %>% select(TeamId = team_id,
                                  FullName = name_last_first,
                                  Position = primary_position,
                                  PosGroup = position_desig,
                                  Bats = bats,
                                  Throws = throws,
                                  Number = jersey_number,
                                  PlayerID = player_id)
            # Finally we return the roster
          return(roster)}
    }
          
# Here we use do.call and apply to call the function
# This will run the rosterFunction above on each row in allTeamIds (which is a 30x1 d.f. with each team id)
# and ultimately rbind each output together.
AllPlayers <-  do.call(rbind,
                       apply(allTeamIds,
                             1, # applying the function on EACH row
                             rosterFunction)
)

# Here we cerate the final tibble
# We join back to the allTeams d.f. to map teamId to teamName 
FinalPlayers <- as_tibble(AllPlayers) %>% inner_join(allTeams %>% 
               select(name_display_long, team_id),
               by = join_by("TeamId" == "team_id")) %>%
  rename("TeamName" = "name_display_long")

return(FinalPlayers)
}
          
        
```
`Player Batting Statistics`  
This function will take a player and return their batting statistics for a given season. The playerName should be given in Last, First format.

```{r}

playerBattingStatistics <- function(playerName, year){
 
 ### This function is built to return a player's statistics during a season ###
 ### Inputs:
 ### playerName -- a Major League Baseball player's full name in Last, First format, i.e. "Harper, Bryce"
 ### team -- a Major League Baseball team's full display name, i.e. "Philadelphia Phillies"
 ### Ex. playerStatistics("Arozarena, Randy", 2022) will return Arozarena's statistics
 ### during the 2022 regular season. 

   # If a year is not given properly (i.e. not numeric) then 2023 will be the given value.
  if(is.numeric(year)){
    year <- year
  }else{
    year <- 2023
  }
  
  # If a player is not given properly (i.e. not a character) then Bryce Harper will be the given value.
  if(is.character(playerName)){
      playerName <- playerName
    }else{
      playerName <- "Harper, Bryce"
    }
  
  # First we call our LeagueRosters function to get all players for that year
AllPlayers  <- LeagueRosters(year)
# Then we filter down to our playerID of interest
PlayerId <- AllPlayers %>% filter(FullName ==  playerName) %>% select(PlayerID)
# Here is the final extraction step for the playerID
PlayerIdFinal <- PlayerId$PlayerID
         
# Here we read in the JSON from the hitting stats endpoint
statsRead <- 
  GET(paste0("http://lookup-service-prod.mlb.com/json/named.sport_hitting_tm.bam?league_list_id='mlb'&game_type='R'&season='"
      , year,"'&player_id='"
      , PlayerIdFinal, "'"))

# We parse that data and get the data.frame object of interest
statsParsed <- fromJSON(rawToChar(statsRead$content))
# We bind all of the relevant hitting statistics together 
statsBind <- cbind(
  # Here we rename some statistics and convert them all to numeric 
   Season =  as.numeric(statsParsed$sport_hitting_tm$queryResults$row$season),
   League = statsParsed$sport_hitting_tm$queryResults$row$league,
   Team = statsParsed$sport_hitting_tm$queryResults$row$team_full,
   PlayerId =  as.numeric(statsParsed$sport_hitting_tm$queryResults$row$player_id),
   Games = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$g),
   PA = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$tpa),
   HR = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$hr),
   RBI = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$rbi),
   AVG =  as.numeric(statsParsed$sport_hitting_tm$queryResults$row$avg),
   OBP = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$obp),
   SLG = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$slg),
   OPS = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$ops),
   BABIP = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$babip),
   SB = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$sb),
   BB = as.numeric(statsParsed$sport_hitting_tm$queryResults$row$bb)
)
# We build the binded stats as a tibble, and also add the player's name
StatsTibble <- as_tibble(statsBind) %>% 
  mutate(PlayerName = playerName) %>% select(PlayerName, everything())

# Finally we return the stats tibble
return(StatsTibble)
}

```

`Player Pitching Statistics`  
This function will take a player and return their pitching statistics for a given season. The playerName should be given in Last, First format.

```{r}

playerPitchingStatistics <- function(playerName, year){
 
 ### This function is built to return a player's statistics during a season ###
 ### Inputs:
 ### playerName -- a Major League Baseball player's full name in Last, First format, i.e. "Kershaw, Clayton"
 ### team -- a Major League Baseball team's full display name, i.e. "Philadelphia Phillies"
 ### Ex. playerStatistics("Kershaw, Clayton", 2019) will return Kershaw's statistics
 ### during the 2019 regular season. 

  
   # If a year is not given properly (i.e. not numeric) then 2023 will be the given value.
  if(is.numeric(year)){
    year <- year
  }else{
    year <- 2023
  }
  
  # If a player is not given properly (i.e. not a character) then Clayton Kershaw will be the given value.
  if(is.character(playerName)){
      playerName <- playerName
    }else{
      playerName <- "Kershaw, Clayton"
    }
  
  
# First we call our LeagueRosters function to get all players for that year
AllPlayers  <- LeagueRosters(year)
# Then we filter down to our playerID of interest
PlayerId <- AllPlayers %>% filter(FullName ==  playerName) %>% select(PlayerID)
# Here is the final extraction step for the playerID
PlayerIdFinal <- PlayerId$PlayerID
         
# Here we read in the JSON from the pitching stats endpoint
statsReadPitching <- 
   GET(paste0("http://lookup-service-prod.mlb.com/json/named.sport_pitching_tm.bam?league_list_id='mlb'&game_type='R'&season='"
      , year,"'&player_id='"
      , PlayerIdFinal, "'"))

# We parse that data and get the data.frame object of interest
statsParsed <- fromJSON(rawToChar(statsReadPitching$content))
# We bind all of the relevant pitching statistics together 
statsBind <- cbind(
  # We rename and convert statistics to numeric
  Season = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$season),
  League = statsParsed$sport_pitching_tm$queryResults$row$league,
  TeamName = statsParsed$sport_pitching_tm$queryResults$row$team_full,
  Wins = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$w),
  Losses = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$l),
  Games = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$g),
  GamesStarted = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$gs),
  IP = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$ip),
  ERA = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$era),
  WHIP = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$whip),
  SOper9 = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$k9),
  HitPer9 = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$h9),
  CG = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$cg),
  SHO = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$sho),
  SV = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$sv),
  SVO = as.numeric(statsParsed$sport_pitching_tm$queryResults$row$svo)
)
# We build the binded stats as a tibble, and also add the player's name
StatsTibble <- as_tibble(statsBind) %>% 
  mutate(PlayerName = playerName) %>% select(PlayerName, everything())

# Finally we return the stats tibble
return(StatsTibble)
}

```

`All Player Stats`  
Here we're going to combine some of the functions above into one and retrieve player statistics for each player on a team in a given season.

I realized that it's rare to search for just one player. Getting statistics for all players will lead to a better analysis with more observations. "All" can be used to specify all teams within that season, or individual teams can be listed out in a vector. This function will return it's final object as a list. The first list will be hitters and the second list will be pitchers.

```{r}

AllPlayerStats <- function(year, teamNames){
  
 ### This function is built to return a multiple player's statistics during a season ###
 ### Inputs:
 ### year -- a season of interest, i.e. 2023
 ### team -- a vector of a Major League Baseball team's full display name, i.e. "Philadelphia Phillies"
 ### Ex. AllPlayerStats(2023, c("Los Angeles Dodgers", "Texas Rangers")) will return both pitching and hitting statistics for each player on those teams.

  
   # If a year is not given properly (i.e. not numeric) then 2023 will be the given value.
  if(is.numeric(year)){
    year <- year
  }else{
    year <- 2023
  }
  # If a year is not given properly (i.e. not a character) then "All" will be the given value.
  if(is.character(teamNames)){
    teamNames <- teamNames
  }else{
    teamNames <- "All"
  }
  
 # Retrieving all players using our LeagueRosters function 
 AllPlayers  <- LeagueRosters(year)
 # Filtering down to teams that were input
 if("All" %in% teamNames){
   # If the input is All then get AllPlayers
   TeamPlayers <- AllPlayers
 
   }else{
   # If the input has specified teams use those! 
     
     TeamPlayers <- AllPlayers %>% filter(TeamName %in% teamNames)
 
     }
 
 # Filtering down on all pitchers
 PitcherNames <- TeamPlayers %>% filter(Position == "P") %>% select(FullName)
  
 # Acquiring pitcher Statistics
 pitchersDf <- data.frame() # setting blank df
 for(i in 1:nrow(PitcherNames)){
   # Using tryCatch to catch all errors. The loop will still continue despite error.
   # That row will juts get skipped.
  pitcherStats <- try(expr = playerPitchingStatistics(PitcherNames[i,]$FullName, year),
                      silent = TRUE)
  
  # Storing each pitcher's statistics in this DF
  pitchersDf <- rbind(pitcherStats, pitchersDf)
  
 }

 # Filtering down on non-Pitchers
 HitterNames <- TeamPlayers %>% filter(Position != "P") %>% select(FullName)
  
 # Acquiring hitter Statistics
 hittersDf <- data.frame() # setting blank df
 for(i in 1:nrow(HitterNames)){
   # Using try to catch all errors. The loop will still continue despite error.
   # That row will juts get skipped.
  hitterStats <- try(expr = playerBattingStatistics(HitterNames[i,]$FullName, year),
                      silent = TRUE)
  # Storing each hitter's statistics
  hittersDf <- rbind(hitterStats, hittersDf)
  
 }
 
 # Joining back to AllPlayers to get positions
 hittersFinal <- hittersDf %>% inner_join(TeamPlayers %>% select(FullName, Position),
                          join_by("PlayerName" == "FullName"))
 
 pitchersFinal <- pitchersDf %>% inner_join(TeamPlayers %>% select(FullName, Position),
                          join_by("PlayerName" == "FullName"))
 
 # Returning a list of both data frames
   return(list(hittersFinal, pitchersFinal))

}

```

# Exploratory Data Analysis

Since we're finished up writing functions to interact with the API, let's put them to use and get some data to conduct analysis!

Baseball is full of statistics. The study of statistics has been recently applied to baseball exhaustively, thanks to a sabermetric onslaught in the early 2000s led by the Oakland Athletics, who followed Bill James' statistical ideas behind the game. This was eventually adapted into the film Moneyball. 

OBP, or On-Base-Percentage, is a controversial statistic in today's game of baseball. While it tells you how often a player gets it wasn't a valued part of a player's game before the sabermetrics caught the league by storm. Now, it can be a million-dollar difference maker in a contract for a player or team. After all teams can only score runs if players get on base.

The four best teams in terms of Runs Scored were the Atlanta Braves, Los Angeles Dodgers, Texas Rangers, and Tampa Bay Rays. These same four teams led the league in OBP. Let's use our API function to get their player statistics. This function might take a moment to run. 

```{r}
# Calling our function and storing it in the object
EDA_TeamStatistics <- 
  AllPlayerStats(2023, c("Atlanta Braves", "Los Angeles Dodgers", "Texas Rangers", "Tampa Bay Rays"))

```

Now that we have the stats, we can specifically grab the batter stats and look into OBP by team. Considering they were the four best at scoring runs, I'd think to see a large number of above average players per team, with the Braves' having the most (as they finished first in Runs). The league average OBP was .320 -- let's see how many players each team had above, at, and below that threshold. 

```{r}
# Splitting our object to get only team data
TeamHittingData <- EDA_TeamStatistics[[1]]

# Creating a categorical variable for above or below league average
ContigencyTableOneData <- TeamHittingData %>% mutate(
  OBP_Level = case_when(
    as.numeric(OBP) < .320  ~ "Below Average",
    as.numeric(OBP) == .320 ~ "Average",
    as.numeric(OBP) > .320  ~ "Above Average"))

# Printing contigency table
table(ContigencyTableOneData$Team, ContigencyTableOneData$OBP_Level)

```
Despite scoring more runs than both the Rays and the Rangers, the Dodgers had the smallest number of players with an above average OBP. However, they also had the least number of below average players. I am a little shocked to see the Rays with the highest number of above average players, considering they finished fourth out of these teams in runs scored.

We can also view the number of players who had an above average, average, or below average SLG% -- which essentially measures the total number of bases per at bat, giving more credit to Home Runs and extra-base-hits than singles. Using similar code above, we get another contingency table:

```{r}
# Creating a categorical variable for above or below league average
ContigencyTableOneData <- TeamHittingData %>% mutate(
  SLG_Level = case_when(
    as.numeric(SLG) < .414  ~ "Below Average",
    as.numeric(SLG) == .414 ~ "Average",
    as.numeric(SLG) > .414  ~ "Above Average"),
  )

# Printing contigency table
table(ContigencyTableOneData$Team, ContigencyTableOneData$SLG_Level)

```
The Braves, who were first in runs scored, had the highest number of players with above average SLG%. This makes sense -- a higher SLG% means a higher number of total bases per at-bats, which in turns means more extra-base-hits and more runs. They also tied the Dodgers for the least number of players with a below average SLG%. 

Two of these teams are front running the MVP race in the National League. The Atlanta Braves hold two of the four favorites in Ronald Acuna Jr. and Matt Olson, while the Los Angeles Dodgers hold the other pair of top candidates in Mookie Betts and Freddie Freeman -- a former Atlanta Brave himself. 

We have already seen that the four of these teams are very good offensively. I think it would be hard to say that one of these players are necessarily more valuable to their teams than the others, given how good each team is overall. However, I'd like to determine which player *is* technically more "valuable" compared to other players, as well as their teams/position groups in order to make a final decision. I will be focusing on six statistics: Home Runs (HR), Runs Batted In (RBI), Stolen Bases (SB), Walks (BB), and Slugging Percentage (SLG%), and On-Base Percentage (OBP%).

Let's compare these four players and retrieve their statistics. 
```{r}
# Running our function for each player
playerName <- "Acuna Jr., Ronald"
Acuna <- playerBattingStatistics(playerName, 2023)

playerName <- "Olson, Matt"
Olson <- playerBattingStatistics(playerName, 2023)

playerName <- "Betts, Mookie"
Betts <- playerBattingStatistics(playerName, 2023)

playerName <- "Freeman, Freddie"
Freeman <- playerBattingStatistics(playerName, 2023)

# Putting these players together
MVPs <- rbind(Acuna, Olson, Betts, Freeman)

```

First, we can look at a simple bar chart for Home Runs. 
If you are of the mindset that Home Runs should determine who wins the award, then Matt Olson is your MVP! He led all of Major League Baseball with 54 Home Runs this year.

```{r}
g <- ggplot(MVPs, aes(x = PlayerId, y= HR))
g + 
  geom_bar(aes(fill = PlayerId), stat = "identity") +
  labs(x = "Player", y = "Home Run Total") +
  scale_x_discrete(labels = c("Freedie Freeman",
                              "Mookie Betts",
                              "Matt Olson",
                              "Ronald Acuna Jr."
                              )) +
  ggtitle("Home Runs By Player") +
  theme_bw()


```
Unfortunately for Olson, there is a lot more at stake to be named the MVP in my book, albeit this is a strong point for Olson's case. With the data we have, I'd like to look at how these players compare to others on their teams across the other offensive statistics. Freeman is in last place here by a longshot, which doesn't bode well for him unless he stands out in other areas.

Next, let's try to display a player's skillset of power and speed. We can represent power with SLG% and speed with SB.

```{r}

# Filtering down all 4 team statistics to Dodgers and Braves
FilteredTeamStats <- EDA_TeamStatistics[[1]] %>% filter(Team %in% c("Atlanta Braves", "Los Angeles Dodgers")) %>% mutate(
 SLG_Level = case_when(
    as.numeric(SLG) < .414  ~ "Below Average",
    as.numeric(SLG) == .414 ~ "Average",
    as.numeric(SLG) > .414  ~ "Above Average"))

# Base plot
g <- ggplot(FilteredTeamStats, aes(x = SLG_Level, y = as.numeric(SB)))
# Using Faceting
g + 
  geom_point() +
  # Display graph for each grouping
  facet_wrap(~ Team) + 
  # Labels
  labs(x = "SLG Level", y = "Stolen Bases") +
  # Adding specific colors for our players of interest
  geom_point(color = ifelse(FilteredTeamStats$PlayerName %in%
                                            c("Acuna Jr., Ronald",
                                            "Olson, Matt",
                                            "Betts, Mookie",
                                            "Freeman, Freddie"), "green", "black")) +
  ggtitle("SB vs. SLG% By Team") +
  theme_bw() 

```
As mentioned before, SLG% is the number of total bases per at-bat for a player. A higher SLG% indicates a player with more "power", or Extra Base Hits (hits that are not singles). It's noticeable that each player has an above average SLG% but Ronald Acuna Jr. (the uppermost left point) stands far and above the rest thanks to his 73 stolen bases. A power-speed combination of that level is rare, and makes Acuna a unique player. This is a strong lobbying point for Acuna.

Another important factor in considering which player is most valuable would be to consider these players by their position grouping. Olson and Freeman are both infielders. Betts and Acuna Jr. are both outfielders. How much better are these players compared to those in their position groups? *We are excluding Deisignated Hitters for this plot.*

```{r}

# Filtering the data and mutating a new column
BoxPlotData <- FilteredTeamStats %>%  mutate(
 PosGroup = case_when(
    Position %in% c("LF", "CF", "RF")  ~ "Outfield",
    Position %in% c("C", "1B", "2B", "3B", "SS") ~ "Infield",
    Position %in% c("DH") ~ "DH"))

par(mfrow = c(2, 2)) # Create a 2 x 2 plotting matrix

# Each of these four plots follow the same technique
g <- ggplot(BoxPlotData %>% filter(PosGroup != "DH"), aes(x=Team, y=as.numeric(HR))) # setting vars.
g + geom_boxplot(aes(fill = PosGroup)) + # filling by pos. group
  theme_bw() +  # theme
  # Title and labels
  ggtitle("Box Plot of HR By Pos. Group By Team") +
  labs(x = "Teams", y = "Home Runs") 

g2 <- ggplot(BoxPlotData %>% filter(PosGroup != "DH"), aes(x=Team, y=as.numeric(SB)))
g2 + geom_boxplot(aes(fill = PosGroup)) +
  theme_bw() + 
  ggtitle("Box Plot of SB By Pos. Group By Team") +
  labs(x = "Teams", y = "Stolen Bases")

g3 <- ggplot(BoxPlotData %>% filter(PosGroup != "DH"), aes(x=Team, y=as.numeric(RBI)))
g3 + geom_boxplot(aes(fill = PosGroup)) +
  theme_bw() + 
  ggtitle("Box Plot of RBI By Pos. Group By Team") +
  labs(x = "Teams", y = "RBI")

g4 <- ggplot(BoxPlotData %>% filter(PosGroup != "DH"), aes(x=Team, y=as.numeric(BB)))
g4 + geom_boxplot(aes(fill = PosGroup)) +
  theme_bw() + 
  ggtitle("Box Plot of BB By Pos. Group By Team") +
  labs(x = "Teams", y = "BB")

```
When box plotting the four statistics above, we see that Home Runs are skewed right. Stolen Bases are approximately normal. RBIs and BBs are relatively normal across the board as well, but skewed right for the Atlanta Braves infield group.

The only outlier points on these graphs belonging to any of the four MVP candidates are Acuna's. His 80 walks and 73 stolen bases lead Atlanta's Outfield group by a wide-margin. Freeman, Betts, and Olson do not appear as outliers on any of these charts. Acuna is the only player who is an outlier for their team and group out of these four players.

What if we considered just position groups across these two teams? We can consider numerical summaries of the offensive statistics of interest.
```{r}
# Creating a new table from the BoxPlot data (ignoring DH's again)
# Summarizing across many variables
SummaryNumAVG <- BoxPlotData %>% filter(PosGroup != "DH") %>% 
  group_by(PosGroup) %>%
  summarize("Min.AVG" = min(as.numeric(AVG)),
            "1stQ AVG" = quantile(as.numeric(AVG), 0.25),
            "Median AVG" = quantile(as.numeric(AVG), 0.5),
            "Mean AVG" = mean(as.numeric(AVG)),
            "3rdQ AVG" = quantile(as.numeric(AVG), 0.75),
            "Max AVG" = max(as.numeric(AVG)),
            "SD AVG" = sd(as.numeric(AVG)))
            

# Creating a new table from the BoxPlot data (ignoring DH's again)
# Summarizing across many variables
SummaryNumOBP <- BoxPlotData %>% filter(PosGroup != "DH") %>% 
  group_by(PosGroup) %>%
          summarize("Min.OBP" = min(as.numeric(OBP)),
            "1stQ OBP" = quantile(as.numeric(OBP), 0.25),
            "Median OBP" = quantile(as.numeric(OBP), 0.5),
            "Mean OBP" = mean(as.numeric(OBP)),
            "3rdQ OBP" = quantile(as.numeric(OBP), 0.75),
            "Max OBP" = max(as.numeric(OBP)),
            "SD OBP" = sd(as.numeric(OBP)))


# Creating a new table from the BoxPlot data (ignoring DH's again)
# Summarizing across many variables
SummaryNumSLG <- BoxPlotData %>% filter(PosGroup != "DH") %>% 
  group_by(PosGroup) %>%
          summarize("Min.SLG" = min(as.numeric(SLG)),
            "1stQ SLG" = quantile(as.numeric(SLG), 0.25),
            "Median SLG" = quantile(as.numeric(SLG), 0.5),
            "Mean SLG" = mean(as.numeric(SLG)),
            "3rdQ SLG" = quantile(as.numeric(SLG), 0.75),
            "Max SLG" = max(as.numeric(SLG)),
            "SD SLG" = sd(as.numeric(SLG))
            )

# Calling each summary and player stats
SummaryNumAVG
SummaryNumOBP
SummaryNumSLG
MVPs %>% select(PlayerName, AVG, OBP, SLG)

```
Even when combining both teams, Acuna stands out well past the 3rd Quantile in AVG, OBP, and SLG for his position group. Betts also stands out, but is not as far ahead as Acuna. Olson is just above his potion group's 3rd Quantile for AVG, but well above in OBP and SLG. Freeman is comfortably ahead in each of these categories as well, but, none of them as far as Acuna.

We'll take into account one more detail and statistic -- the idea of luck. BABIP, Batting Average on Balls put In Play, can signify a player being "lucky." If a player has a high BABIP, that generally signifies some level of luck, albeit, not perfectly quantifiable. This could indicate getting a really weak hit ground ball past a fielder because they got a poor reaction, as opposed to really hitting a ball hard and getting into the Outfield. My guess would be that BABIP and AVG/OBP are correlated. This plot below comes from the ggcorrplot() package.
```{r}
# Creating new DF for this plot
PercentageStats <- EDA_TeamStatistics[[1]] %>% select(AVG,
                                                      OBP,
                                                      SLG,
                                                      BABIP,
                                                      HR,
                                                      RBI,
                                                      BB,
                                                      SB)
# Cleaning the data and making them all numeric
corData <- mutate_all(PercentageStats, function(x) as.numeric(as.character(x)))

# Using the ggcorrplot package
# Creating matrix as that is the input for the ggcorrplot package
p.mat <- cor_pmat(corData)

# Plotting
ggcorrplot(p.mat)

```
Shockingly, BABIP and each percentage variable, OBP, SLG, and AVG, are uncorrelated. However, BABIP is correlated strongly with BBs and HRs. Let's take a look at a numerical summary for BABIP and see how players stack up by position group.
```{r}

# Creating our summary for the position group
SummaryFiveNumBABIP <- BoxPlotData %>% filter(PosGroup != "DH") %>% 
  group_by(PosGroup) %>%
          summarize("Min. BABIP" = min(as.numeric(BABIP)),
            "1stQ BABIP" = quantile(as.numeric(BABIP), 0.25),
            "Median BABIP" = quantile(as.numeric(BABIP), 0.5),
            "Mean BABIP" = mean(as.numeric(BABIP)),
            "3rdQ BABIP" = quantile(as.numeric(BABIP), 0.75),
            "Max BABIP" = max(as.numeric(BABIP)),
            "SD BABIP" = sd(as.numeric(BABIP)))

SummaryFiveNumBABIP
MVPs %>% select(PlayerName, BABIP)

```
Olson, Acuna, and Betts are all within about one standard deviation of the mean BABIP. This doesn't say much about them being lucky or unlucky. However, Freeman is nearly four standard deviations past the average BABIP for a his position group among the Ddogers and Braves. I would consider him to be more "lucky" than the other players. 

If Freeman we're the second horse in this race, this would strain his campaign, but I think it's clear to see that Acuna should be the MVP. His speed-power combination is something that is rare in the game of baseball, he ranks second among candidates in Home Runs, and ranks first in AVG, OBP, and OPS. He also stands out within his position group for his team, more than the other players do so for their team's respective position groups, perfectly depicting his value.

# Wrap Up
Thanks for taking the time to read through this project! I would've like the MLB API endpoint to house more team-related information like record during the season, winning percentage, and such. It would've been fun to dig into that data here. Please refer to my github blog for a longer reflection on this project.

# Creating .md File
```{r}
# 
# rmarkdown::render("~/repos/558_project2/Matarazzo_Project.rmd", 
#               output_options = list(
#                 name_value_pairs = "value", 
#                 toc = TRUE,
#                 toc_depth = 2,
#                 number_sections = TRUE,
#                 df_print = "tibble",
#                 or_something = TRUE
#                 ),
#                 output_file = "Matarazzo_Output"
#               )
```

